---
title: "Possum Final Project "
author: Jake Truman, Isaih Barron, and Will Martin
output: pdf_document
---

```{=html}
<style type="text/css">
h1.title {
font-size: 40px;
text-align: center;
}
h4.author {
font-size: 40px;
text-align: center;
}
</style>
```
\newpage

```{r setup, include=FALSE}
# load all necessary packages here
library(bestglm)
library(dplyr)
library(gridExtra)
library(corrplot)
library(caret)
library(ggplot2)
library(ggfortify)
library(car)
library(patchwork)
```

# Abstract

Our data set involves the measurements of a species of possum over seven sites in Australia, ranging from Southern Victoria to Queensland, which encompasses almost all of Australia's eastern coast. In this analysis, our goal is to create a model capable of accurately predicting the total length of a possum based on key characteristics such as location, sex, age, etc. With this, we will be able to analyze trends in the data, which may reveal how the possums in specific areas interact with their environment.

## 1.1 Data Description

It is important to sample animal populations as it allows us to monitor the status of a species. This information lends to important conclusions that may explain trends in an ecosystem's health. Biologists are able to use the information collected from animals to asses the behavioral patterns of animals, as well as the dynamics of the population. Ultimately, this possum data set aids in the management and protection of these possum and their habitats. Down below is a table of relevant data. In this analysis, the relevant variables are *case, site, Pop, hdlngth, totlngth, taill*.

| Variable | Description                                                                |
|--------------------------|----------------------------------------------|
| case     | Observation Number                                                         |
| site     | The site number where the possum was trapped                               |
| Pop      | Population: either Vic (Victoria) or other (New South Wales or Queensland) |
| hdlngth  | Head length in millimeters (mm)                                            |
| totlngth | Total length in centimeters (cm)                                           |
| taill    | Tail length in centimeters (cm)                                            |

## 1.2 Questions of Interest

For this data set, there are two questions of interest:

Question 1: Is there a difference between using a the site a possum was captured compared to the population the possum belongs to.

Question 2: Does the site where a possum was captured have an effect on the total length of the possum.

## 1.3 Regression Methods

For the questions, to determine if and how the site of a possum's capture has an effect on the total length of the possum, a multiple linear regression model was fit using the *site* a possum was caught at and the *chest* *size*, *head length*, and *tail length o*f a possum as the predictor variables in addition to another model with the same variables except *site* is swapped with the population the possum is from. The best subset variable selection method with the AIC metric was used to determined which variables to build the model on. Since *site* is categorical variable, it was encoded as a factor.

After some exploratory data analysis, an interesting item found (in regards to the first question) are the box plots relating the population and site a possum is from to the same possum's total length. As shown in the box plots for populations, there is not much of a difference between the different populations; however, some of the box plots for site are noticeably different, like sites two and four. This suggests that using *site* may be more useful. (Additionally, a summary for the data is given in Figure 11. )

```{r}
# Read and Format the Data
data = read.csv("possum.csv", sep =",")
data = data.frame(data)
data = na.omit(data)
data = data[,-1]
data$Pop = factor(data$Pop, levels = c("other","Vic"))
data$sex = factor(data$sex, levels = c("m","f"))
data$site = factor(data$site, levels = c("1","2","3","4","5","6","7","8","9","10"))
data = data %>% relocate(totlngth, .after = belly)

y=(data$sex)
X=as.matrix(data[, 2:13])

pop_plot <- ggplot(data, aes(x = Pop, y = totlngth)) +
  geom_boxplot()

site_plot <- ggplot(data, aes(x = site, y = totlngth)) +
  geom_boxplot()

(pop_plot) /
(site_plot)
```

# 2 Analyses, Results, and Interpretation

## Diagnostic Checks

In order to ensure that our first model created above isreliable, the model assumptions -- linearity, independence, normality of residuals, homoscedasticity, no multicollinearity , and no influential points -- need to be checked.

The first assumption checked will be that there is a linear relationship between the response variable (*total length* of the possum) and the predictor variables. Looking at the added-variable plots below, there is a somewhat strong linear relationship between each of the predictor variables and the response except for one predictor: *site 3*. Also looking at the residuals vs. fitted-values plot, each of the points are centered and normally distributed around zero. Although the one predictor with *site 3* is somewhat worrisome, for simplicity this assumption is likely met.

Considering the independence between observations, not much information was disclosed about how the data was collected; however, what was said is this data came from ecological research for statistical analysis. Because of this, likely the data was collected in a professional way to ensure independence, so this assumption will likely be met.

Next is the assumption that the residuals of the model are normally distributed. Looking at the histogram (see Figure 2) of residuals and the Shapiro Test, the histogram is normally distributed and the null hypothesis that the residuals are normally distributed is not rejected. This assumption is easily met.

For the equal and constant variance assumption, looking at the scale-location plot, the line is somewhat straight and the points are evenly distributed across it. Additionally, looking at the Q-Q plot, most of points follow the diagonal line closely except for the lower theoretical quantiles. This is potentially concerning, but as most values follow diagonal line and the scale-location plot looks mostly good, this is assumption is likely met

Next is checking that there is no multicollinearity among the predictor variables. The correlation matrix (see Figure 8) and VIFs were computed (see Figure 3). In the correlation matrix down, none of the variables have a higher correlation than `0.8`, which is ideal, and the mean and max of the VIFs was `1.90` and `2.89` respectively, which is less ideal but likely still good enough. The assumption is likely met and will not interfere much with the predictions of the model.

Finally is the check for influential points. Looking at both residuals vs. leverage plots (see Figure 1 and the plot below), all data points were within bounds (less than `0.5`), meaning that none of them could be considered influential.

```{r, echo = FALSE}
# Fit the Model
multReg = lm(totlngth ~ site + chest + hdlngth + taill, data = data)
data$Residuals = multReg$residuals
data$fitted = multReg$fitted.values

# Added-Variable Plots (for Linearity)
avPlots(multReg)

# Normality of Residuals
shapiro.test(data$Residuals)

# LINE Assumptions and Influential Points
autoplot(multReg)
```

## Analysis Details and Interpretation

To answer the first question about our data set, an analysis and other tests on the models were performed using 2 different models with variations on those models. These two models are the Site model and Population model. The Site model was a linear model that used *site*, *chest*, *head length* and *tail length* as predictor variables with the response variable being Total Length (See Figure 4). The Population model was a similar linear model that used *chest, population*, *tail length*, and *head length* as predictors and Total Length as the response (See Figure 6).

First, an anova test was conducted to see if the inclusion of the *population* and *site* variables were producing statically significant different models. (In other words, checking to see how the addition of a categorical variable, *population* or *site*, affected the model.) For the models that did and did not contain *site*, the anova analysis produced a p-value of $7.604e^{-11}$ suggesting that the inclusion of *site* in the model (along with *chest*, *head* *length*, and *tail* *length*) compared to a model without site has a statistically difference (See Figure 5). Next, the inclusion of *population* in the Population model was tested. The anova analysis resulted in a p-value of $2.228^{-8}$, thus indicating that a model with *chest, tail length* and *head length*, *population* is a statically significant difference compared to a model without *population* (See Figure 7).

To further understand the relationship that site and population have as predictor variables in a linear model, we conducted an anova test comparing those two models (see Figure 9). The anova analysis resulted in a p-value of $3.145^{-5}$ which indicates that the difference between models is statically significant at an alpha level of 0.05.

Now that we have compared the models, we can look at how they actually are structured. For this section, we will compare the predictor variables that both models share (Tail Length and Head Length), what effect they have, and compare the difference categorical variables make in the models. This will answer the second question about our data set.

Starting with the Site model, the estimated coefficients for the predictor variables are given in Figure 4. The tail length coefficient is interpreted as: holding all else constant we expect that on average for every `1` centimeter increase in tail length we expect the total length to increase by `1.19337` centimeters. For interpreting the coefficient for head length: holding all else constant, we expect that on average for every `1` millimeter increase in head length we expect the total length to increase by `0.40183` centimeters. We now give a generalization about the site factor: because all of the coefficients of determination for each site are negative, we acknowledge that each site has a significant effect on the total length of a possum. We note that, holding all else constant, the site factor will always decrease the total length of the possum in comparison to the reference Site (Site 1). We note specifically that Site 3 has the smallest effect on Total Length and that Site 7 has the largest effect.

With the Population model, the estimated coefficients for the predictor variables are given in Figure 6. The tail length is interpreted as: holding all else constant, we expect that for every `1` centimeter increase in tail length, the total length will increase by `1.31775` centimeters, on average. For head length, holding all else constant, we expect that for every `1` millimeter increase in head length, the total length will increase by `0.49654` centimeters, on average. For the categorical variable, it can be interpreted as: holding all else constant, a possum belonging to the Vic population is expected to have a total length `3.27880` centimeters greater than a possum from the "Other" population, on average.

Finally, we created 95% confidence intervals for both models as shown in Figure 10. This was done to further determine the effect of predictor variables.

We started with the Site model, which has a categorical variable with 6 levels. Once again, we made a generalization of the results. With all of the site confidence intervals being negative, and all of them spanning a similar distance, we can assume, holding all else constant, with 95% confidence that the true value for each of the sites is within the upper and lower bound of the confidence intervals (see Figure 10). We also see that none of the confidence intervals contain 0, which allows us to infer that the site variable will always result in a decrease in the total length of a possum. We can interpret the confidence interval for tail length as: we are 95% confident that, holding all else constant, the true effect of Tail Length on Total Length lies between `0.94229399` and `1.444456`. Finally, we are 95% confident that, holding all else constant, the true effect of chest on total length lies between `-0.07118978` and `0.4579026`. Since the confidence interval contains 0, we recognize that Chest might not have any statistically significant effect on Total Length, at the significance level of `.05` (see Figure 10).

We now look at the Population model. We are 95% confident that, holding all else constant, the true effect of tail length on total length is between `1.0799613` and `1.6158617`. Finally, we are 95% confident that, holding all else constant, the true effect of a possum being part of the Vic Population on the total length is between `2.5141664` and `4.5601573`. We did not include confidence intervals for Head Length in the discussion, due to Head Length being a predictor variable with a relatively low effect on both models. The intervals are included in Figure 10 in the appendix.

# 3 Conclusions

The analyses we conducted were done from the scope of statisticians, so we will put it in simpler terms. We had two problems we wanted to answer. The First: does the site a possum was captured at to determine the total length of the possum? The Second: When explaining the total length of a possum, is there a difference between the sub population of a possum and the site a possum was captured at? To answer the first question, we used a method of linear regression that was able to explain roughly 80% of the variation in a possums length when we use the site of its capture. We tackled the second question in a similar fashion, and concluded that there is a difference in using the sub population compared to the Site of a possums capture when explaining the variation in total length. As this was a nonembellished version of the conclusion, we will explain the analysis more in depth in the following paragraphs.

We started with an anova test between models that either had or didn't have Site as a categorical predictor variable. The test resulted in a p-value well below the significance level of `0.05` which meant they were statistically significantly different. Comparing the model summaries, we saw that when Site was included we had an $R^2 =.80$ compared to the other model which returned $R^2 = .64$ . Including the categorical variable of site allowed for us to explain roughly 16% more of the data. Additionally, we compared models where the categorical predictor variable was Site and Population by performing an anova test which resulted in a statistically significant difference. The Population model had an $R^2 =.74$ which was less than the aforementioned value of the Site model. We also compared the Site model's $Adjusted\ R^2 =.79$ with the Population's $Adjusted\ R^2 =.73$ . Using both those metrics the Site model was able to explain more of the data while still being better suited to generalize potential new data. We believe the reason for this difference is related to the Sites in which possums were captured. If a site had more food or a harsher environment, it would make sense that the possums would be larger there, due to abundance of nutrients or the need to defend from predators. However, since we have no additional data we are merely speculating on the potential reason why Site as a predictor variable allowed models to perform better.

There are a few issues we felt may be potentially concerning. First, we wanted to try testing an interaction term between Population and Site, but were unable to do so as the multicollinearity between the two variables was far too high. Also, when constructing confidence intervals for the coefficients of the Site model, zero was included in the interval for the Chest variable. This typically indicates that the variable is not statistically significant in the model, but since the interval was only 0.07 below zero we determined that it was acceptable to keep it in our model.

In summary, our study explored the influencing factors of a possums Total Length, specifically examining the capture site and sub population variables. Our analysis, using multiple linear regression models, uncovered insight into the relationship between these variables. The implications of the findings suggest that the environment and conditions of a possum significantly effect the animal's total length.

# 4 Contributions

Jake was responsible for conducting the anova test between models and creating the confidence intervals. Jake also performed the variable selection methods and construction of the models. In addition, Jake helped with making sure the assumptions were met and performed data cleaning.

Isaih did the main checking and analysis of the assumptions, smoothing out of the writing, and helped come up with some potential questions. He also helped explore some early models that are not shown here.

Will helped with the first regression model and aided in the analysis. He was also tasked with constructing the introductory paragraphs that introduce the dataset as well as the questions of interest. Towards the end of the project, he was in charge of editing in order to improve the cohesion of the report.

## APPENDIX

### Figure 1

```{r,echo=FALSE}
cd_cont_pos <- function(leverage, level, model) {sqrt(level*length(coef(model))*(1-leverage)/leverage)}
cd_cont_neg <- function(leverage, level, model) {-cd_cont_pos(leverage, level, model)}

cd_threshold <- 0.5
autoplot(multReg, which = 5) +
  stat_function(fun = cd_cont_pos,
                args = list(level = cd_threshold, model = multReg),
                xlim = c(0, 0.6), lty = 2, colour = "red") +
  stat_function(fun = cd_cont_neg,
                args = list(level = cd_threshold, model = multReg),
                xlim = c(0, 0.6), lty = 2, colour = "red") +
  scale_y_continuous(limits = c(-4, 4))
```

### Figure 2

```{r,echo=FALSE}
ggplot(data = data) +
  geom_histogram(aes(x = Residuals, y = after_stat(density)), 
                 binwidth = 0.5) +
  stat_function(fun = dnorm, color = "red", linewidth = 2,
                args = list(mean = mean(data$Residuals), 
                            sd = sd(data$Residuals))) 
```

### Figure 3

```{r,echo=FALSE}
vif_values <- car::vif(multReg)
cat("Mean VIF:",mean(vif_values))
print("Individual VIFS")
print(vif_values)
```

### Figure 4

```{r,echo=FALSE}
site_model = lm(totlngth ~ site + chest + hdlngth + taill, data = data)
summary(site_model)
```

### Figure 5

```{r,echo=FALSE}
no_site_model = lm(totlngth ~ chest + hdlngth + taill , data = data)
anova(site_model, no_site_model)
```

### Figure 6

```{r,echo=FALSE}
pop_model = lm(totlngth ~ chest + hdlngth + taill + Pop, data = data)
summary(pop_model)
```

### Figure 7

```{r,echo=FALSE}
no_pop_model = lm(totlngth ~ chest +hdlngth+ taill, data = data)
anova(no_pop_model, pop_model)
```

### Figure 8

```{r,echo=FALSE}
# Exlude site as that is not numeric
data_subset <- data[c('totlngth', 'chest', 'hdlngth', 'taill')] 
corrplot(cor(data_subset))


```

### Figure 9

```{r,echo=FALSE}

anova(site_model, pop_model)
```

### Figure 10

```{r,echo=FALSE}
print("Population Model Confidecnce Interval")
print(confint(pop_model, level =0.95))
print("Site Model Confidecnce Interval")
print(confint(site_model, level =0.95))
```

### Figure 11

```{r}
summary(data)
```
